---
title: "Assingment NYC Flight Data"
author: "Your Name Here"
date: "October 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, `airports.csv`, `planes.csv` and `weather.csv`.  Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
df <- read.csv("~/csx460/CSX460/02-fundamentals/data/flights.csv")
```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
hist(df$dep_delay, main = "Departure delay", xlab = "Time")
hist(df$arr_delay, main = "Arrival delay", xlab = "Time")

```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}

plot(df$time_hour, df$arr_delay, main = "Date vs Arrival Delay", xlab = "Date", ylab = "Arrival Delay")

Or

DateMDY <- as.Date(paste(df$year, df$month, df$day), format = "%Y%m%d")
df$Date <- paste(DateMDY)
plot(df$Date, df$arr_delay, main = "Date vs Arrival Delay", xlab = "Date", ylab = "Arrival Delay")
```



## Create tables for Categorical Variables

Create a `table` of occurences of origin airports and carrier.

```{r}

CrossTable(df$origin, df$carrier)

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}

```



