---
title: "Assingment NYC Flight Data"
author: "Your Name Here"
date: "October 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, `airports.csv`, `planes.csv` and `weather.csv`.  Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}
df = read_csv("~/csx460/CSX460/02-fundamentals/data/flights.csv")
```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}
hist(df$dep_delay, main = "Departure delay", xlab = "Time")
hist(df$arr_delay, main = "Arrival delay", xlab = "Time")

```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}

plot(df$time_hour, df$arr_delay, main = "Date vs Arrival Delay", xlab = "Date", ylab = "Arrival Delay")

Or

DateMDY <- as.Date(paste(df$year, df$month, df$day), format = "%Y%m%d")
df$Date <- paste(DateMDY)
plot(df$Date, df$arr_delay, main = "Date vs Arrival Delay", xlab = "Date", ylab = "Arrival Delay")
```



## Create tables for Categorical Variables

Create a `table` of occurences of origin airports and carrier.

```{r}

CrossTable(df$origin, df$carrier)

```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}
df_weather <- read_csv("~/csx460/CSX460/02-fundamentals/data/weather.csv")
df_planes <- read_csv("~/csx460/CSX460/02-fundamentals/data/planes.csv")
df_planes <- read_csv("~/csx460/CSX460/02-fundamentals/data/planes.csv")


df_join_df_flight_name <- df %>% left_join(df_planes, by = "tailnum")

df_join_df_flight_name_join_weather2 = df_join_df_flight_name %>% left_join(df_weather, c("month", "day", "hour", "origin"), c("year.x" = "year"))

df_join_df_flight_name_join_weather2_join_origin_airports = df_join_df_flight_name_join_weather2 %>% left_join(df_airports, c("origin" = "faa"))

df_join_df_flight_name_join_weather2_join_origin_airports_by_dest_airport = df_join_df_flight_name_join_weather2_join_origin_airports %>% left_join(df_airports, c("dest" = "faa"))
```



