---
title: "NYCFlights: Arrival Delay Regression Model"
author: "Snehal A Saraf"
date: "17 October 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("ggplot2")
install.packages("psych")
install.packages("lubridate")
```

```{r load_libs}
library(ggplot2)
library(psych)
library(readr)
library(magrittr)
library(lubridate)
library(stringr)
library(dplyr)
library(data.table)
```

#Exercise 1
```{r "Solution to exercise 1"}
setwd("~/csx460/CSX460/02-fundamentals")
flights  <- "data/flights.csv" %>% read_csv 
planes   <- read_csv("data/planes.csv")
airports <- read_csv("data/airports.csv") 
weather  <- read_csv("data/weather.csv")

str_pad4 <- . %>% str_pad(4,'left',0)
str_pad2 <- . %>% str_pad(2,'left',0)

ymd_hm <- function(year, month, day, hour, minute ) { 
  dates <- paste0(
    year %>% str_pad4, 
    month %>% str_pad2,
    day %>% str_pad2,
    " ",
    hour %>% str_pad2,
    ":",
    minute %>% str_pad2
  ) 

  dates %>% lubridate:::ymd_hm(.)

}

ymd_hm1 <- function(year, month, day, hour, minute ) { 
  dates <- paste0(
    str_pad(4, 'left', year),
    str_pad(2,'left', month),
    str_pad(2,'left', day),
    " ",
    str_pad(2,'left', hour),
    ":",
    str_pad(2,'left',minute)
  ) 

  dates %>% lubridate:::ymd_hm(.)

}

flights <- flights %>% mutate( dt = ymd_hm(year,month,day,hour,minute) )

flights$carrier %>% table( useNA="always") %>% sort 
flights$origin %>% table( useNA="always") %>% sort 

table( flights$carrier, flights$origin)

YX <- 
  flights %>% 
  left_join(weather, by=c("origin","year","month","day","hour") ) %>%
  left_join(planes, by='tailnum', suffixes=c('','.pl') ) %>% 
  left_join(airports, by=c('origin'='faa') ) %>% 
  left_join(airports, by=c('dest'='faa') )    

```

# NYCFlights Model

Using the rectangular data that you created last week and following the *Predicting Medical Expenses* example from the text (*MLwR*), create a model for arr_delay. Follow MLwR structure 
for building a model. Describe/Explain each of the steps and show all work in codeblocks below.


## Step 1: Collect Data 

```{r "Step 1: Collect Data" }
str(YX)
```

## Step 2: Explore and Prep The Data


#One of the things not done in the MLwR text is a pairwise comparison between the response and each of the predictors. Make sure to do this; this is often very illustrative of the relationship between that predictor and the response. This can be done with `pairs` or `psych::panel.pairs`


```{r "Step 2: Explore and Prep The Data" }
summary(YX$arr_delay)

hist(YX$arr_delay)

ggplot(data=YX, aes(x=arr_delay)) + geom_histogram()

ggplot(data=YX, aes(x=log(10, YX$arr_delay))) + geom_histogram()


ggplot(data=YX, aes(x=arr_delay)) + geom_histogram() + scale_x_sqrt()

ggplot(data=YX, aes(x=arr_delay)) + geom_histogram() + scale_x_log10("Arrival Delay") + scale_y_continuous("Count")

smp <- sample(1:nrow(YX), size = round(0.7*nrow(YX)), replace=FALSE)
train <- YX[smp,]
test <- YX[-smp,]
cor(train[c("arr_delay", "dep_delay", "dep_time", "sched_arr_time", "distance", "air_time", "temp", "wind_speed", "wind_gust", "wind_dir", "precip", "visib")], use="complete")
#pairs(YX[c("arr_delay", "dep_delay", "dep_time", "sched_arr_time", "distance", "air_time", "temp", "wind_speed", "wind_gust", "wind_dir", "precip", "visib")])
```

## Step 3:  Train The Model

```{r "Step 3: Train The Model" }


# Throws error ->
m1 <- lm(arr_delay ~ ., data = train)

m1 <- lm(arr_delay ~ dt + dep_time + sched_dep_time + sched_arr_time + arr_time + carrier + flight + origin + dest + air_time + distance + dewp + temp + wind_dir + wind_speed + wind_gust + visib + pressure + precip + engines + seats + speed + type + manufacturer + model + engine, data = train)


```

## Step 4: Evaluate Performance

Evaluating performance is a crutial step, that has yet to be treated well by the MLwR. Think about some of the tools that you have for evaluating performance.  Choose one and articulate why you have chosen it.

```{r "Step 4: Evaluate Performance" }
mpredict <- predict(m1, test)

#Plot graph of the errors.
```


## Step 5: Improve Performance 

#features Not needed  -
# tailnum 
# names <airport names> since we have the codes.
# Manufacturer since we already have model

# Combination features -
# dep_time * sched_dep_time
# 

Show some steps for improving model performance.

```{r "Step 5: Improve Performance" }

```


# Question:

Is this a good model?  (Write your answer here.)

