---
title: "Improving Model Perfromance / Tuning Parameters"
author: "Snehal Ashok Saraf"
date: "`r Sys.Date()`"
output: html_document
---


## Tuning Parameter

Generically and regardless of model type, what are the purposes of a model
tuning parameters?

```
It is used to tune the model to enhance the output of the model. 
```

## Caret Models

This assignment demonstrates the use of caret for constructing models. Each
model should be built and compared using using `Kappa` as the performance
metric calculated using 10-fold repeated cross-validation with 3 folds.

Using the rectangular data that you created for the NYCFlights to create a model
for arr_delay >= 15 minutes.

- glm
- rpart
- knn
- C50
- randomForest
- adaBoost
- Two methods of your choice from the Caret Model List (you will need to install any dependencies)

Save the caret objects with the names provided.

```{r load_libs}
library(caret)
library(ggplot2)
library(psych)
library(readr)
library(magrittr)
library(lubridate)
library(stringr)
library(dplyr)
```

```{r}

# Your work here.

# Read the data
#df <- read_csv("~/csx460/CSX460/02-fundamentals/data/flights.csv") 
load("~/csx460/CSX460/02-fundamentals/data/ex2.RData")
setwd("~/csx460/CSX460/02-fundamentals/data/")

# Study the table given and cherry pick the important features

y <- "arr_delay"
xs <- c(
    'month','dep_delay','carrier','air_time','distance'
   ,'year.y','type','engine'
   ,'wind_dir','wind_speed','wind_gust','precip','pressure','visib'
   ,'lat.x','lon.x','lat.y','lon.y'
)

df <-  YX[, c(y, xs)]
df$delayed <- ifelse(df$arr_delay>=15, 1, 0)
str(df)
# Create a new df with {classification = 0 if arr_delay < 15 mins} and

# Plot graph of o/p
ggplot(data = df, aes(delayed)) + geom_histogram()
df$arr_delay <- NULL


#Get NA count
df %>% sapply(. %>% is.na %>% sum)
library("Amelia")
missmap(df)

# year.y, engine, type has lots of NA's.
df$year.y <- NULL
df$type <- NULL
df$engine <- NULL

# delayed itself has NA . Delete rows with NA's for delayed.
df <- df[!is.na(df$delayed),]

# Impute the NA's in each column. To avoid error while running train -
# Error in na.fail.default(list(delayed = c(0, 1, 1, 0, 0, 0, 1, 0, 0, 0,  :
# Refer https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/
# Not needed since na.pass in train for caret takes care of imputation.

#library(mice)
#mids <- mice(df, m= 1, maxit = 1, method = "pmm", seed = 52)
#summary(mids)
#midsbkup <- mids
#mids$imp$dep_delay
# Get the first iteration
#df <- complete(mids, 1)

set.seed(1244)

#Create a train control to indicate that "k fold repeated cv" data prep is to be used
# selectionFunction is used if the grid param is to be used which allows us to control
# the value of the params to the "train". It means out of the various models built using
# various values of the parameter, HOW do we choose one ? <eg: best one OR next to best one etc >

# 10 fold cv with 3 repetitions
trainctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3, selectionFunction = "oneSE")


# For classification, we need the outcome "delayed" to be a factor rather than a numeral. Else
# train considers it as a regression
df$delayed <- ifelse(df$delayed == 1, '1', '0')

# na.action = na.pass imputes the NA's automatically for us. However, it does NOT work for trControl
# when a fold-cv is used. Hence, use na.action = na.omit
fit.glm <- train(delayed ~ ., 
                 data = df, 
                 method = "glm",
                 na.action = na.omit, 
                 trControl = trainctrl, 
                 metric = "Kappa",
                 family = "binomial")

fit.knn <- train(delayed ~ ., data = df, method = "knn", na.action = na.pass)
fit.rpart <- train(delayed ~ ., data = df, method = "rpart", na.action = na.pass)
fit.rf <- train(delayed ~ ., data = df, method = "rf", na.action = na.pass)
fit.myown1 <- ..
fit.myown1 <- ..

fit.glm
```

Compare the  models?

Which is best?  Why?

```
```
